Transmembrane helices are an overlooked and evolutionarily conserved source of major histocompatibility complex class I and II epitopes

### 2021-11-30 bbbq_article.tex
Article refer to figures ascendingly
Process reviewer feedback, thanks @fransbianchi and @bogeert

This study had two objectives. 
First, we aimed to generalize our findings by predicting
the antigenic presentation
from different kingdoms of life in both MHC-I and -II. 
From these \emph{in silico} predictions, we conclude that TMH-derived
epitopes from a human, viral and bacterial proteome
are likely to be presented more often than expected by chance 
for most alleles of MHC-I and II. 
We confirmed the presentation of TMH-derived peptides 
by re-analysis of peptides from The Immune Epitope Database (IEDB) \cite{vita2019immune}.
Second, we tested our hypothesis that TMHs 
are more evolutionary conserved than soluble protein regions.
Our analysis of human single nucleotide polymorphisms (SNPs) showed 
that random point mutations are indeed less likely
to occur within TMHs. 
These findings strengthen the emerging notion 
that TMHs are important for the T cell-mediated adaptive immune system, 
and hence are of  
overlooked importance in vaccine development.

@@

Epitope prediction is important to understand the immune system function
and for the design of vaccines.
In this study, we provide evidence that epitopes
derived from TMHs are a major but overlooked source of MHC epitopes. 
Our bioinformatics predictions indicate that the TMH-derived epitope repertoire is larger than expected by chance for both MHC-I and MHC-II, regardless of the organism. Moreover, reanalysis of MHC-ligands from the IEDB database confirmed the presentation of TMH-derived epitopes. Therefore, it seems likely that TMH-derived epitopes would also result in enhanced T cell responses, although the conservation of TMHs might promote the deletion of T cells responsive to TMH-derived epitopes by central tolerance mechanisms. Finally, our SNP analysis shows that TMHs are evolutionary more conserved 
than solvent-exposed protein regions.

@@

  \caption{
    \textbf{
      Robust prediction that TMH epitopes are presented.
    }
    The percentage of epitopes for MHC-I and -II alleles that overlap with TMHs
    that are presented. The pair of horizontal red lines in each plot indicate the
    lower and upper bound of the 99\% confidence interval. Note that only one line
    is visible as this interval is relatively narrow. 
    Alleles are listed in Table \ref{tab:haplotype_abbreviations}).
    \textbf{(A)} 
    Observed and predicted percentage of TMH-derived epitopes for MHC-I alleles. 
    \textbf{(B)} 
    MHC ligands from IEDB corresponding to TMH-derived epitopes. 
    The numbers above the bars denotes the number of TMH derived epitopes obtained.
  }
  \label{fig:elution}
\end{figure}

>>>>

This study had two objectives. 
First, we aimed to generalize our findings by predicting
the antigenic presentation
from different kingdoms of life in both MHC-I and -II. 
From these \emph{in silico} predictions, we conclude that TMH-derived
epitopes from a human, viral and bacterial proteome
are likely to be presented more often than expected by chance 
for most alleles of MHC-I and II. 
We confirmed the presentation of TMH-derived peptides 
by re-analysis of peptides from The Immune Epitope Database (IEDB) \cite{vita2019immune}.
Second, we tested our hypothesis that TMHs 
are more evolutionary conserved than soluble protein regions.
Our analysis of human single nucleotide polymorphisms (SNPs) showed 
that random point mutations are indeed less likely
to occur within TMHs. 
These findings strengthen the emerging notion 
that TMHs are important for the T cell-mediated adaptive immune system, 
and hence are of importance in vaccine development.

@@

Epitope prediction is important to understand the immune system function
and for the design of vaccines.
In this study, we provide evidence that epitopes
derived from TMHs are a major source of MHC epitopes. 
Our bioinformatics predictions indicate that the TMH-derived epitope repertoire is larger than expected by chance for both MHC-I and MHC-II, regardless of the organism. Moreover, reanalysis of MHC-ligands from the IEDB database confirmed the presentation of TMH-derived epitopes. Therefore, it seems likely that TMH-derived epitopes would also result in enhanced T cell responses, although the conservation of TMHs might promote the deletion of T cells responsive to TMH-derived epitopes by central tolerance mechanisms. Finally, our SNP analysis shows that TMHs are evolutionary more conserved 
than solvent-exposed protein regions.

@@

  \caption{
    \textbf{
      Analysis of epitope database shows that TMH derived epitopes are over presented.
    }
    The percentage of epitopes for MHC-I and -II alleles that overlap with TMHs
    that are presented. The pair of horizontal red lines in each plot indicate the
    lower and upper bound of the 99\% confidence interval. Note that only one line
    is visible as this interval is relatively narrow. 
    Alleles are listed in Table \ref{tab:haplotype_abbreviations}).
    \textbf{(A)} 
    Observed and predicted percentage of TMH-derived epitopes for MHC-I alleles. 
    \textbf{(B)} 
    MHC ligands from IEDB corresponding to TMH-derived epitopes. 
    The numbers above the bars denotes the number of TMH derived epitopes obtained.
  }
  \label{fig:elution}
\end{figure}
